Elephant Seal Age JAE Intraspecific Variation

### 2023-11-09 paper/paperv2.qmd
Fixing the plot labels.

library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)

@@

seal_phenology <- sealdat %>% 
  mutate(ba = firstobsbreeddoy, 
         bd = firstobsbreeddoy + breeddur, 
         ma = lastobsmoltdoy - moltdur, 
         md = lastobsmoltdoy) %>% 
  drop_na(age, ba, bd, ma, md)

@@

timing_summary <- seal_phenology %>% 
  group_by(age, observed) %>% 
  summarize(across(ba:md, mean), 
            .groups = "drop") %>% 
  pivot_longer(cols = -c(age, observed),  
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age, observed) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Long trip", 
                                          "Breeding", 
                                          "Short trip", 
                                          "Molting", 
                                          "longtrip2")))
timing_summary_flipped <- timing_summary %>% 
  mutate(observed = ifelse(observed == "Breeder",
                           "Non-breeder",
                           "Breeder")) %>% 
  filter(event != "longtrip2")

@@

# This data frame contains a model for each phenological variable w.r.t. age
phenology_models <- seal_phenology %>% 
  pivot_longer(ba:md, names_to = "phenology", values_to = "doy") %>% 
  select(phenology, observed, doy, age, animalID, year) %>% 
  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
  nest(model_data = c(doy, age, observed, animalID, year)) %>% 
  mutate(model = map(model_data, fit_model),
         age_coef = extract_estimate(model, "age"),
         age_stderr = extract_stderr(model, "age"),
         age_pval = extract_pval(model, "age"),
         age_signif = age_pval * 8 < 0.05,
         breeder_coef = extract_estimate(model, "observedBreeder"),
         breeder_stderr = extract_stderr(model, "observedBreeder"),
         breeder_pval = extract_pval(model, "observedBreeder"),
         breeder_signif = breeder_pval * 8 < 0.05,)

@@

phen_plot2 <- phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(c(ends_with("coef"), ends_with("stderr"), ends_with("signif")),
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(signif = as.logical(signif)) %>% 
  ggplot(aes(x = coef, y = phenology)) +
  # Prevent geom_rect() from establishing a continuous y-axis scale
  geom_point(alpha = 0) +
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = event_begin, ymax = event_end,
                fill = event),
            phen_wrapped,
            inherit.aes = FALSE) +
  geom_pointrange(aes(xmin = coef - stderr, 
                      xmax = coef + stderr,
                      color = signif)) +
  scale_color_manual(values = c("FALSE" = "grey50", "TRUE" = "black"),
                     guide = "none") +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  expand_limits(x = 0) +
  facet_grid(cols = vars(param), scales = "free_x") +
  theme_few()

phen_plot2

library(patchwork)
phen_plot1 /
  phen_plot2 +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(legend.position = "top",
        legend.title = element_blank())

plot_grid(phen_plot1, phen_plot2, nrow = 2, labels = "AUTO", align = "v")

@@

## Discussion

>>>>

library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)
library(patchwork)

@@

seal_phenology <- sealdat %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD)

@@

timing_summary <- seal_phenology %>% 
  group_by(age, observed) %>% 
  summarize(across(BA:MD, mean), 
            .groups = "drop") %>% 
  pivot_longer(cols = -c(age, observed),  
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age, observed) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Long trip", 
                                          "Breeding", 
                                          "Short trip", 
                                          "Molting", 
                                          "longtrip2")))
timing_summary_flipped <- timing_summary %>% 
  mutate(observed = ifelse(observed == "Breeder",
                           "Non-breeder",
                           "Breeder")) %>% 
  filter(event != "longtrip2")

@@

# This data frame contains a model for each phenological variable w.r.t. age
phenology_models <- seal_phenology %>% 
  pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
  select(phenology, observed, doy, age, animalID, year) %>% 
  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
  nest(model_data = c(doy, age, observed, animalID, year)) %>% 
  mutate(model = map(model_data, fit_model),
         age_coef = extract_estimate(model, "age"),
         age_stderr = extract_stderr(model, "age"),
         age_pval = extract_pval(model, "age"),
         age_signif = age_pval * 8 < 0.05,
         breeder_coef = extract_estimate(model, "observedBreeder"),
         breeder_stderr = extract_stderr(model, "observedBreeder"),
         breeder_pval = extract_pval(model, "observedBreeder"),
         breeder_signif = breeder_pval * 8 < 0.05,)

@@

phen_plot2 <- phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(c(ends_with("coef"), ends_with("stderr"), ends_with("signif")),
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(signif = as.logical(signif), 
         param = factor(param, 
                        labels = c("Maternal Age (Years)", 
                                   "Breeding Status"))) %>% 
  ggplot(aes(x = coef, y = phenology)) +
  # Prevent geom_rect() from establishing a continuous y-axis scale
  geom_point(alpha = 0) +
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = event_begin, ymax = event_end,
                fill = event),
            phen_wrapped,
            inherit.aes = FALSE) +
  geom_pointrange(aes(xmin = coef - stderr, 
                      xmax = coef + stderr,
                      color = signif)) +
  scale_color_manual(values = c("FALSE" = "grey50", "TRUE" = "black"),
                     guide = "none") +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  expand_limits(x = 0) +
  facet_grid(cols = vars(param), scales = "free_x") +
  xlab("LMM coefficeint") + 
  ylab("Phenology") + 
  theme_few()

phen_plot1 /
  phen_plot2 +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(legend.position = "top",
        legend.title = element_blank())

@@
